---
title: "Regression Analysis (Econ 0211), Lab1 : Introduction"
author: "Noe Vyizigiro"
date: "March, 2024"
output:
  html_document:
    theme: cerulean
    toc: TRUE #include table of content
    toc_flaot: TRUE
    
---

# *Important packages/libraries for this lab*:

1. tidyverse
2. foreign: used when reading data that is not in .CSV format
3. readstata13 : used when reading data in dta format

```{r}

library(tidyverse) 
if(!require("tufte")) install.packages("tufte")
#library(tufte)

if(!require("foreign")) install.packages("foreign")
library(foreign) #this will help when importing data that are not in csv.format(stata, sas, spss)
#library(haven)

if(!require("readstata13")) install.packages("readstata13")
library(readstata13)

rm(list = ls()) #clear the environment

```
# I. Loading the data set in R. 
*We use function "rad.dta13" from the library "readstata13" which read the dta version of dataset from stata version 13 and above*
```{r}
Cps20 <- read.dta13("CPS_2020_monthly.dta")
Cps21<-read.dta13("CPS_2021_23_monthly.dta")

```

# III. Looking at the data: Describe, Browse, Summarize, Tabulate
### 7.
```{r, eval=FALSE}
#quickly look at data structure
glimpse(Cps20) 

#Summary statistic of all variables in the dataset
summary(Cps20) 

prop.table(table(Cps20$month)) # to check the frequency distribution of the value of a variable of interest; shows proportion.
table(Cps20$month)#gives number of occurrence

head(Cps20$month) # see a specific variable from the dataset

is.na(Cps20$unemployed) # check missing data from a variable (return a true or false)
sum(is.na(Cps20$unemployed)) # find count of missing values
which(is.na(Cps20$unemployed)) #find the location of missing variable
```

### 8. Counting number of occurence of a variable
```{r}
# tabulate month
Cps20 %>%
  count(month)%>%# we use count function
  mutate(Percent=round(n/sum(n),4)*100)%>% # frequency rounded at 3
  arrange(desc(Percent))%>% # arranged in descending order
  mutate(Percentile = cumsum(Percent))


```
# IV. Generating new variables; use mutate to add a new variable to the dataset
### 10. 
```{r}
# tab to view dummy variable unemployed created
Cps20%>%
  group_by(empstat)%>% #group by empstat
  summarise(mean_unemployed = mean(unemployed),
            n =n()) # compute the mean of unemployed
```

### 11. create a dummy variable female
```{r}
Cps20<-Cps20 %>%
  mutate(female=ifelse(sex=='female',1,0))

```

### 12. Compute the mean of the female variable
```{r}
# tabulate female
Cps20%>%
  count(female) %>%
  mutate(percent = round(n/sum(n),4)*100)

# summarise female
Cps20%>%
  summarise(Obs =n(),
            mean_female = round(mean(female),3),
            Std.dev. =round(sd(female),3),
            min = min(female),
            max = max(female))

# optional: calculating average unemployment rate by sex in each month
Cps20 %>% 
  group_by(month,sex) %>%
  summarise(aver_unemployed =round(mean(unemployed, na.rm =TRUE),3))%>%
  pivot_wider(names_from = sex,
              values_from = aver_unemployed)%>%
  ungroup()
```

### 13. Generate a dummy variable 'college' for at least having 4-year college degree.

*Create a dummy variable "college", which is equal to 1 if educ ( a person) has a bachelor's degree, master's degree, or a doctorate degree*
*we use case_when function when we are treating more than one condition ( using more than one if-statement)*
```{r}

Cps20<-Cps20 %>%
  mutate(college =case_when(Cps20$educ=="bachelor's degree"~1,
                           Cps20$educ=="master's degree"~1,
                           Cps20$educ=="doctorate degree"~1,
                           TRUE~0))  
```

# V. Creating Tables and Figures to describe economic trends

### 14. Compute unemployment rate accross months

In stata, it's only one line of code: tab month, sum(unemployed)
```{r}

# number of employment and unemployment in each month: 
table(Cps20$month, Cps20$unemployed)# tab month unemployed.

# or you can do it in this way:
unemployment<-Cps20%>%
  filter(!is.na(month)& !is.na(unemployed))%>%
  count(month, unemployed)%>%
  pivot_wider(names_from = unemployed,
              values_from = n) %>%
  rename(unemployment_1="1",
         employment_0="0")
unemployment


#Mean of unemployment rates in each month 
Cps20%>%
  group_by(month)%>%
  #filter(month=="january")%>%
  summarise(Obs = n(),
            mean_unemployed = round(mean(unemployed, na.rm = TRUE), 3),
            Std.dev. = round(sd(unemployed, na.rm=TRUE),3),
            min =min(unemployed, na.rm = TRUE),
            max =max(unemployed, na.rm = TRUE))
 #*To compute the mean at a specific month, use filter function.* 
```

### 15. Unemployment Across Gender

```{r}
sum_female<-Cps20%>% 
  group_by(month,female) %>%
  summarise(aver_unemployed = mean(unemployed, na.rm =TRUE))%>%
  pivot_wider(names_from = female,
              values_from = aver_unemployed)%>%
  rename(unemp_male ="0",
         unemp_female="1")
sum_female
```

### 16. Plots average unemployment accross gender
*Plotting using R is optional,You can plot that in excel by using sum_female*
```{r}
ggplot(sum_female)+
  geom_line(aes(month, unemp_male, group=1, color="male" ))+
  geom_line(aes(month, unemp_female, group=1, color="female"))+
  labs(
    title = "Average unemployment rate accross gender",
    subtitle = "in 2020",
    x="Month",y="gender",
    color=""
  )
  
```

### 17. Generate age categories from age variable

```{r, include=FALSE}
 Cps20_1<-Cps20%>% 
  mutate(age =as.numeric(age))%>%
  mutate(agecat = case_when(
    age >= 20 & age < 30 ~ "1",
    age >= 30 & age < 40 ~ "2",
    age >= 40 & age < 50 ~ "3",
    age >= 50 & age < 60 ~ "4",
    age >= 60 & age < 70 ~ "5",
    age >= 70 & age < 80 ~ "6",
    age >= 80 & age < 90 ~ "7",
    age >= 90 & age <= 150 ~ "8",
    TRUE ~ "0"))
#Cps20_1
```
### Unemployment accross age group
```{r}
Cps20_1%>%
  group_by(month, agecat)%>%
  summarise(aver_unemp=mean(unemployed, na.rm = TRUE))
```

### Graph unemployment accross age group
```{r}
age_graph<-Cps20_1%>%
  group_by(month, agecat)%>%
  summarise(aver_unemp=mean(unemployed, na.rm = TRUE))%>%
  pivot_wider(names_from = agecat,
              values_from = aver_unemp)%>%
  rename(cat0 ="0", cat1="1", cat2="2",cat3="3", cat4="4", cat5="5")%>%
  
  ggplot()+
  geom_line(aes(month, cat0, group=1, color="age<20"))+
  geom_line(aes(month, cat1, group=1, color="between age 20 &30"))+
  geom_line(aes(month, cat2, group=1, color="between age 30 &40"))+
  geom_line(aes(month, cat3, group=1, color="between age 40 &50"))+
  geom_line(aes(month, cat5, group=1, color="between age 50 &60"))+
  geom_line(aes(month, cat5, group=1, color="age>60"))+
  labs(
    title = "Unemployment rates accross age category: 2020",
    x="Months", y="Unemploymnt_rates",
    color ="Age Groups"
  )+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.1, vjust = 0))

 age_graph 
```

### 18. Merge two datasets: append(merge across observation)
```{r}

Cps20_23<-bind_rows(Cps20, Cps21)
```

### Average unemployment accross years (2020-2023):table

```{r}
Cps20_23%>%
  group_by(year)%>%
  summarise(Obs =n(),
            unemp_yrs=round(mean(unemployed, na.rm = TRUE),3),
            Std.dev. = round(sd(unemployed, na.rm =TRUE),3),
            min = min(unemployed, na.rm = TRUE),
            max = max(unemployed,na.rm = TRUE)
  )
            
```

# Visualizing unemployment accross years(2020-2023)

```{r}
Cps20_23%>%
  group_by(year,month)%>%
  summarise(unemp_yrs=mean(unemployed, na.rm = TRUE))%>%
  pivot_wider(names_from = year, 
              values_from = unemp_yrs)%>%
  rename(yr_2020="2020", yr_2021="2021", yr_2022="2022", yr_2023="2023")%>%
  ggplot()+
  geom_line(aes(month, yr_2020, group=1, color="2020"))+
  geom_line(aes(month, yr_2021, group=1, color="2021"))+
  geom_line(aes(month, yr_2022, group=1, color="2022"))+
  geom_line(aes(month, yr_2023, group=1, color="2023"))+
  labs(
    title = "Unemployment rates",
    y="Unemployment_rates", x="Months",
    color = "Years"
  )+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.1, vjust = 0))

```

#this is the end! thank you.




