---
title: "Regression Analysis (Econ 0211), Lab3: Unit of measurement, functional form, and summy regressors"
author: "Noe Vyizigiro"
date: "March, 2024"
output:
  html_document:
    theme: cerulean
    toc: TRUE #include table of content
    toc_flaot: TRUE
---
#Note: Install all the necessary packages before loading the libraries if they are not installed already
## *Important packages (libraries) for this lab*:

1. tidyverse
2. cowplot

```{r}
library(tidyverse)
library(cowplot) # to put graphs together

rm(list = ls()) #clear the environment
```

# 1. loading data
```{r, include=FALSE}
usa21_inc<-read_csv("IPUMS_USA_ 2021_ACS_LAB_3.csv") 

#load function package "NV.R" that will come in handy when creating summary statistics; use nv_sum() function to get a table of summary statistics
source("/Volumes/middfiles/Classes/Fall23/ECON0211A/Noe_Labs/NV.R")

```
```{r}
nv_sum(usa21_inc, incwage, edyears, earnings, weight = FALSE)
```

# 2. Estimating regression  of earning on years
```{r}
Model1<-lm(earnings~edyears, data = usa21_inc, na.action = na.exclude) # omit N/A's to get predicted values  
summary(Model1)
coef(Model1) # pull out the coefficients
```

# 3. generating residuals and fitted values
```{r}
usa21_inc$yhat=fitted(Model1) # generate fitted values from Model1
#this also works: #usa21_inc$yhat<-Model1$fitted.values 

usa21_inc$uhat<- residuals(Model1) # generate residual values.

cor(usa21_inc$edyears,usa21_inc$earnings) # correlation between two variables

#view the predicted variables
usa21_inc%>%
  na.omit()%>%
  dplyr::select(yhat, uhat) 

```

# creating graphs
```{r}

g1<-ggplot(usa21_inc,aes(x=edyears, y =earnings))+ 
  geom_point()+
  labs(
    title = "Income wage on years of education",
    x="years of education",
    y = " income wage"
  )

# another quick way of plotting is using qplot function: qplot(usa21_inc$edyears,usa21_inc$INCWAGE) +geom_smooth(method = lm, se=FALSE)
g2<-ggplot(usa21_inc,aes(x=edyears, y =earnings))+ # x and y have to be within the ggplot function for the fitted line to work
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(
    title = "Income wage on years of education",
    x="years of education",
    y = " income wage"
  )
 
# plot the residual on fitted values
# the line of best fit is horizontal: there is is not a pattern to this data. So there is a linear relationship.
g3<-ggplot(usa21_inc, aes(x=yhat, y = uhat))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(
    title = "Residual on fitted values",
    x="fitted_values(yhat)",
    y="residuals(uhat)"
  )
plot_grid(g1, g2, g3)
```

# 4. Estimating a regression model of earnings on months of schooling
```{r}
# creating a new variable month
usa21_inc<-usa21_inc%>%
  mutate(edmonths = edyears*10) # assuming each year has 10 months of schooling

# regress earning on months of education
Model2<-lm(earnings~edmonths, data = usa21_inc, na.action = na.exclude)
summary(Model2)

```

# 5. Estimating a regression model of earnings(in cents) on years of education
```{r}
#create a new variable earn_cents
usa21_inc<-usa21_inc%>%
  mutate(earn_cents=earnings*100)

Model3<-lm(earn_cents~edyears, data=usa21_inc)
summary(Model3)
```

# 6. multiplying by 5
```{r}
usa21_inc<-usa21_inc%>%
  mutate(earn_5 = earnings*5,
         ed_5 =edyears*5)
Model4<-lm(earn_5~ed_5, data = usa21_inc)
summary(Model4)
```

# 7. data transformation using log
```{r}
usa21_inc<-usa21_inc%>%
  mutate(ln_earnings=log(earnings))

```
# 8. Estimating a fitted model of log of earnings on years of education.
```{r}
Model5<-lm(ln_earnings~edyears, data = usa21_inc)
summary(Model5)
```

# 9. Dummy variable as explainatory variable
```{r}
# creating a dummy variable for being female and summarize earnings:
usa21_inc%>%
  count(sex)

usa21_inc<-usa21_inc%>%
  mutate(female =ifelse(sex=="female", 1,0))
usa21_inc%>%
  summarise(mean_female=mean(female))

# summarizing using nv_sum()
female_usa21_inc<-usa21_inc%>%
 group_by(female)
nv_sum(female_usa21_inc, earnings, weight = FALSE)

"
#This is the regular summary without using nv_sum function
usa21_inc%>%
  group_by(female)%>%
  summarise(mean_wage =mean(earnings),
            total=n())
"

# Regress earnings on the female
Model6<-lm(earnings~female, data = usa21_inc)
summary(Model6)
```


```{r}
# creating a dummy variable male
usa21_inc<-usa21_inc%>%
  mutate(male =ifelse(sex =="male",1, 0))

#summarize earnings
male_usa21_inc<-usa21_inc%>%
  group_by(male)
nv_sum(male_usa21_inc, earnings, weight = FALSE)
  

# Regress earnings on the male
Model7<-lm(earnings~male, data = usa21_inc)
summary(Model7)

# regress ln_earnings on female
Model8<-lm(ln_earnings~female, data = usa21_inc)
summary(Model8)

```



