---
title: "Regression Analysis (Econ 0211), Lab6: Hypothesis Testing and Motivating Nonlinear Regression"
author: "Noe Vyizigiro"
date: "April, 2024"
output:
  html_document:
    theme: cerulean
    toc: TRUE #include table of content
    toc_flaot: TRUE
---
*This is  a second version of the lab 6 where we use the dataset imported from the stata version of this lab. Using this dataset corrects all the slight mismatch between the r and stata results * 

## *Important packages (libraries) for this lab*:
1. tidyverse: Contain libraries such as dplyr and ggplot, which are instrumental here
2. stargazer: helpful in creating multi-model regression tables
3. sandwich: Helpful in estimating robust regression models
4. lmtest : helpful in hypothesis testing..
5. car : helpful in model testing
6. survey : useful in running weighted statistical tests


#Load libraries
```{r, include=FALSE}
library(tidyverse)

if(!require("stargazer")) install.packages("stargazer")
library(stargazer)

if(!require("sandwich")) install.packages("sandwich") # for robust regression
library(sandwich)

if(!require("lmtest")) install.packages("lmtest")
library(lmtest)

if (!require("survey")) install.packages("survey")
library(survey)

if(!require("car")) install.packages("car")
library(car) # allows different model tests

rm(list = ls())
```

# 1. Read Dataset (.csv file)
```{r}
ATUS<-read.csv("IPUMS_ATUS_2019.csv")

source("/Volumes/middfiles/Classes/Fall23/ECON0211A/Noe_Labs/NV.R") #load NV.R to access nv_sum() function for quick summary statistic table
```

# Generate numeric values for educ
```{r}
ATUS<-ATUS%>%
  mutate(codeduc=with(ATUS, match(educ, unique(educ))))

```
# View the labels created and encode
```{r}
ATUS%>%
  group_by(educ, codeduc)%>%
  summarise(mean_l<-mean(leisure))
#encode
ATUS<-ATUS%>%
  mutate(edyearF=factor(codeduc, c(1,4,5,15,8,2,3,9,13,11,10,6,12,17,14,16,7),
                        labels=c("10","11","11", "3", "6", "7", "9", "14","14","16", "20","12", "12","0", "18","18", "14")),
         edyear=as.numeric(as.character(edyearF)))
#summarize to check the match
ATUS%>%
  group_by(educ, edyear)%>%
  summarise(mean_l = mean(leisure))

 #check employment
ATUS%>%
  group_by(empstat)%>%
  summarise(mean_emp=mean(leisure))

#check kidund18
ATUS%>%
  group_by(kidund18)%>%
  summarise(mean_kid=mean(leisure))
```

# 2. Generate variables
```{r}
ATUS<-ATUS%>%
  mutate(LTHS =ifelse(edyear<12,1,0),
        HS = ifelse(edyear==12,1,0),
        SC = ifelse(edyear>12 & edyear<16, 1, 0),
        BA = ifelse(edyear==16,1,0),
        GRAD=ifelse(edyear>16,1,0),
        employed =ifelse(empstat=="employed - at work",1,0),
        male =ifelse(sex=="male",1,0),
        child=ifelse(kidund18=="yes",1,0))

#summarize the 10 variables; using nv_sum()

nv_sum(ATUS, LTHS, HS, SC, BA, GRAD, employed, male, child, leisure, age, weight = FALSE)
```
# Weighted summary statistics
```{r}
# select variables to summarise
selected_vars <-dplyr::select(ATUS,LTHS, HS,SC, BA, GRAD, employed, male,child, leisure, age)
# Create a survey design object, from survey package
sv_obj <- svydesign(id = ~1, weights = ~wt06, data = ATUS) # id: formula specifying cluster id; ~1 indicate that there is no cluster; weights: formula specifying the sampling weights

# Function to calculate weighted mean and standard error
summary_stats <-function(x,y) { # supply function iterate through the selected variables x while using the weights in the sv_object (y)
  c(
    obs =sum(!is.na(x)),
    mean=svymean(x,y, na.rm=TRUE),#calculate the weighted mean
    Std.dev = sqrt(svyvar(x,y, na.rm = TRUE)),# calculate the weighted SD
    min = min(x, na.rm = TRUE),
    max =max(x, na.rm = TRUE)
   # quant=svyquantile(x,y, 0.50, na.rm=TRUE)
  )
}
sum_table<-sapply(selected_vars, summary_stats, sv_obj) # use the sapply function here to the whole created function to create a vector-type data
#sum_table
# Convert to a data frame for better presentation
results_df <- as.data.frame(t(sum_table))
# Print the results
print(results_df)
```

# 3. Estimating Regression
```{r}
m1<-lm(leisure~HS+ SC+ BA +GRAD, data = ATUS)
#summary(m1)

m2<-lm(leisure~HS+SC+ BA+ GRAD, data =ATUS, weights = wt06) # non robust regression
#summary(m2)

#robust regression
m3<-lm(leisure~HS+SC+ BA+ GRAD, data =ATUS, weights = wt06) 

#rob<-coeftest(m3, vcov = vcovHC(m3, "HC1"))    # robust; HC1 (Stata default)
rob_se3<-sqrt(diag(vcovHC(m3,"HC1"))) #calculate robust standard error
#rob_se3

stargazer( m1,m2,m3,
          se =list(NULL, NULL, rob_se3),
          type = "text",
          title = "Table 1",
          out = "table1.text",
          column.labels = c("Regular", "Weighted", "Robust SE 1"),
          notes = "Significance level"
          
          )
```

# 4. Estimating the regressions using robust option
```{r}
m1<-lm(leisure~HS + SC + BA + GRAD, data =ATUS, weights = wt06)
r_se1<-sqrt(diag(vcovHC(m1,"HC1")))

m2<-lm(leisure~HS + SC + BA + GRAD + employed, data =ATUS, weights = wt06)
r_se2<-sqrt(diag(vcovHC(m2,"HC1")))

m3<-lm(leisure~HS + SC + BA + GRAD + employed + male, data =ATUS, weights = wt06)
r_se3<-sqrt(diag(vcovHC(m3,"HC1")))

m4<-lm(leisure~HS + SC + BA + GRAD + employed + male + child, data =ATUS, weights = wt06)
r_se4<-sqrt(diag(vcovHC(m4,"HC1")))

m5<-lm(leisure~HS + SC + BA + GRAD + employed + male + child + age, data =ATUS, weights = wt06)
r_se5<-sqrt(diag(vcovHC(m5,"HC1")))

stargazer(m1, m2, m3, m4,m5,
          se=list(r_se1, r_se2, r_se3, r_se4, r_se5),
          type = "text",
          title = "Table 2: Weighted Regression with Robust Standard Error.",
          out="table2.txt",
          notes.append = TRUE)
```

# 5. Hypothesis testing
```{r}
#b
m1<-lm(leisure~ HS + SC + BA + GRAD + employed + age + male + child, data = ATUS, weights = wt06)
r_seM1<-sqrt(diag(vcovHC.default(m1, "HC1")))
summary(m1)
r_seM1

t1<-linearHypothesis(m1, "BA =GRAD")
print(t1)
#c
t2<-linearHypothesis(m1, c("HS", "SC", "BA", "GRAD"))
print(t2)
```

# 6. More testing
```{r}
#a
linearHypothesis(m1, c("child =0", "employed=0"))
linearHypothesis(m1, "child-employed=0") # or linearHypothesis(m1, "child=employed =0)
linearHypothesis(m1, "2*child =employed")

#b
linearHypothesis(m1, c("HS", "SC", "BA", "GRAD", "employed", "age", "male", "child "))
```

# 7. Age and Leisure, plots
```{r}
#c
ggplot(ATUS, aes(age, leisure))+
  geom_point()+
  labs(
    title = "scatter plot of age and Leisure"
  )
#Scatter plot is not helpful here, we can try something else
sv_ob <- svydesign(id = ~1, weights = ~wt06, data = ATUS) 

 sum<-ATUS%>%
   group_by(age)%>%
   summarise(mean_leisure = weighted.mean(leisure, wt06),
             n=n())
 #sum
 
 #generate a variable that is the average leisure by age
ATUS <-ATUS%>%
   group_by(age)%>%
   mutate(leisure_byAge = mean(leisure))
  
 ggplot(ATUS, aes(age, leisure_byAge))+
   geom_line()+
   labs(
     title = "Leisure by age and Age"
   )

 # regress leisure on age
 m1<-lm(leisure~age, data = ATUS, weights = wt06)
 summary(m1)
 #robust
 rm1<-coeftest(m1,vcov=vcovHC(m1,"HC1")) #robust coefficient and se
 se_m1<-sqrt(diag(vcovHC(m1, "HC1")))

 
 #atus19$pred_leisure<-predict(m1)
 ATUS$fitted_values<-fitted(m1)
 
ggplot(ATUS, aes(age, leisure_byAge))+
  geom_line(aes(color ="Leisure_byAge"))+
  geom_line(aes(y= fitted_values, color ="fitted_values"))+
   labs(
     title = "Predicted values"
   )
```

# The following steps are ones we will explore in class next week ......
```{r}
#d
ATUS<-ATUS%>%
  mutate(age2 =age^2)

for (i in 2:5){
  new_val_name <- paste0("age", i)
  ATUS[[new_val_name]]<-ATUS$age^i
  
}
```

# Run Regressions and generate predicted values
```{r}
m1<-lm(leisure~ age, weights = wt06, data = ATUS)
ATUS$pred_yht1<-predict(m1)

m1se<-sqrt(diag(vcovHC.default(m1, "HC1")))

m2<-lm(leisure~ age+ age2, weights = wt06, data = ATUS)
ATUS$pred_yht2<-predict(m2)

m2se<-sqrt(diag(vcovHC.default(m2, "HC1")))

m3<-lm(leisure~ age + age2 + age3, weights = wt06, data = ATUS)
ATUS$pred_yht3<-predict(m3)

m3se<-sqrt(diag(vcovHC.default(m3, "HC1")))

m4<-lm(leisure~ age + age2 + age3 + age4, weights = wt06, data = ATUS)
ATUS$pred_yht4<-predict(m4)

m4se<-sqrt(diag(vcovHC.default(m4,"HC1")))

stargazer(m1, m2, m3, m4,
          se =list(m1se, m2se, m3se, m4se),
          type = "text",
          title = "Table 3.",
          out="table3.txt",
          notes.append = TRUE)

```

# Plot the Predicted Values
```{r}
ggplot(ATUS, aes(age, leisure_byAge, color = "leisure_byAge"))+
  geom_line()+
  geom_line(aes(y =pred_yht1, color ="pred_yht1"))+
   labs(
     title = "Predicted values"
   )


ggplot(ATUS, aes(age, leisure_byAge, color = "leisure_byAge"))+
  geom_line()+
  geom_line(aes(y =pred_yht1, color ="pred_yht1"))+
  geom_line(aes(y =pred_yht2, color ="pred_yht2"))+
  geom_line(aes(y =pred_yht3, color ="pred_yht3"))+
  geom_line(aes(y =pred_yht4, color ="pred_yht4"))+
   labs(
     title = "Predicted values"
   )
```

# Final Chart
```{r}
ggplot(ATUS, aes(age, leisure_byAge, linetype = "leisure_byAge", color="leisure_byAge"))+
  geom_line()+
  geom_line(aes(y =pred_yht1, linetype ="linear",color = "linear"))+
  geom_line(aes(y =pred_yht2,linetype ="quadratic",color="quadratic"))+
  geom_line(aes(y =pred_yht3, linetype ="cubic", color="cubic"))+
  geom_line(aes(y =pred_yht4, linetype ="quartic",color="quartic"))+
  scale_linetype_manual(values = c("linear"="dashed", "quadratic"="dotted", "cubic"="twodash", "quartic"="dotdash", "leisure_byAge"="solid"))+
  scale_color_manual(values = c("linear"="blue", "quadratic"="green", "cubic"="purple", "quartic"="red", "leisure_byAge"="black"))+
   labs(
     title = "Predicted values",
     linetype ="Model", # line type will merge with color type to created an annotated, colored legend based on linetype and color type
     color ="Model",
     y="Leisure by Age"
   )+
  theme_minimal()
```




