---
title: "Regression Analysis (Econ 0211), Lab8: Binary Dependent Variables"
author: "Noe Vyizigiro"
date: "April, 2024"
output:
  html_document:
    theme: cerulean
    toc: TRUE #include table of content
    toc_flaot: TRUE
---
# *Important packages/libraries for this lab*:
1. tidyverse
2. stargazer
3. sandwich: robust st.err
4. broom: extract coefficients/margins from linear models
5. lmtest: model tests
6. mfx: computing marginal effects for nonlinear models


```{r, include=FALSE}
library(tidyverse) 

if(!require("stargazer")) install.packages("stargazer")
library(stargazer)

if(!require("mfx")) install.packages("mfx")
library(mfx) #this helps calculate the average partial effect(APE) or the marginal effect

if(!require("sandwich")) install.packages("sandwich")
library(sandwich)

if(!require("lmtest")) install.packages("lmtest")
library(lmtest)
library(margins)

if(!require("broom")) install.packages("broom")
library(broom)


rm(list = ls())
```

# 1. Read data
```{r, include=FALSE}
titanic<-read_csv("titanic.csv")
source("/Volumes/middfiles/Classes/Fall23/ECON0211A/Noe_Labs/NV.R") #load NV.R to access nv_sum() function for quick summary statistic table
```
#2 generate dummy variables

```{r}
titanic<-titanic%>%
  mutate(female =ifelse(gender =="female", 1,0),
         child = ifelse(age<16,1,0),
         survive =ifelse(survived=="Survive", 1,0),
         class1 =ifelse(pclass==1,1,0),
         class2 =ifelse(pclass==2,1,0),
         class3 =ifelse(pclass==3,1,0)
         )
#titanic<-titanic%>%
  #na.omit()

```

# 3. descriptive statistic

```{r}
summary_table<-titanic%>%
  group_by(pclass, gender)%>%
  summarise(mean_survived = mean(survive, na.rm=TRUE),
            )%>%
  pivot_wider(names_from = pclass,
              values_from = mean_survived)
print(summary_table)

```

# 4. Regression
```{r}
m1<-lm(survive~child+female+class1+class2, data=titanic)
m_se1<-sqrt(diag(vcovHC(m1,"HC1"))) #robust se
summary(m1)


```

# 5. Estimate linear probability model
```{r, include=TRUE}
m2<-lm(survive~child+female+class1+class2, data=titanic, na.action = na.exclude)
m_se2<-sqrt(diag(vcovHC(m2,"HC1"))) #robust se

stargazer(m2,
          se =list(m_se2),
          type = "text",
          out = "LPM.txt",
          title = "LPM",
          notes.append = TRUE)

```

# 6. Check predicted outcome
```{r}
summary(m2)

titanic$yhat<-predict(m2) # predict yhat
nv_sum(titanic, yhat, weight=FALSE) # summarise yhat

```

# 7. Probit Regression Model
```{r, include=TRUE}
p_model<-glm(survive~child+female+class1+class2,family = binomial(link = "probit"), data = titanic)
#summary(p_model)
logLik(p_model) # log Likelihood value
lrtest(p_model) # test of overall significance of the model, including the log likelihood value

l_model<-glm(survive~child+female+class1+class2,family = binomial(link = "logit"), data = titanic)
#summary(l_model)
logLik(l_model) # log Likelihood value

stargazer(p_model, l_model,
          type = "text",
          out = "probability.txt",
          title = "Probability Model",
          notes.append = TRUE
  
)
```

# Computing margins/Partial Effects
*To calculate the margins, we use the function logitmfx for a logit model and probitmfx for a probit model. Use tidy function from broom package to extract marginal effects for linear probability model*

```{r}
#linear
lpm<-lm(survive~child+female+class1+class2, data=titanic, na.action = na.exclude)
lpm_tidy<-tidy(lpm)
lme <- lpm_tidy$estimate #extract margins (coefficients)
lse <- lpm_tidy$std.error #extract st err

margins(lpm) # coefficients of the lpm


#probit
pm<-glm(survive~child+female+class1+class2,family = binomial(link = "probit"), data = titanic) # probability model

atmp<-probitmfx(survive~child+female+class1+class2, data = titanic, atmean = TRUE) # Marginal Effect at the average/ Partial Effect Average (PEA)
atmep<-atmp$mfxest[,"dF/dx"] #margins
atse<-atmp$mfxest[,"Std. Err."] #st.errors

avmp<-probitmfx(survive~child+female+class1+class2, data = titanic, atmean = FALSE) # Average Marginal effect /Average Partial Effect (APE)
avmep<-avmp$mfxest[,"dF/dx"] #margins
avse<-avmp$mfxest[,"Std. Err."] #st.error

#logit
lm<-glm(survive~child+female+class1+class2,family = binomial(link = "logit"), data = titanic) # logit model

atml<-logitmfx(survive~child+female+class1+class2, data = titanic, atmean = TRUE) # Marginal Effect at the average/ Partial Effect Average (PEA)
atmel<-atml$mfxest[,"dF/dx"] #margins
atsel<-atml$mfxest[,"Std. Err."] # st.error

avml<-logitmfx(survive~child+female+class1+class2, data = titanic, atmean = FALSE) # Average Marginal effect /Average Partial Effect (APE)
avmel<-avml$mfxest[,"dF/dx"] # margins
avsel<-avml$mfxest[,"Std. Err."] # st.errors


stargazer(lpm,lpm, pm,pm,pm,lm,lm,lm,
          coef = list(NULL,lme,NULL, atmep, avmep,NULL, atmel, avmel),
          se = list(NULL,lse, NULL, atse,avse,NULL,atsel,avsel),
          type = "text",
          out = "probability.txt",
          title = "Probability Model",
          notes.append = TRUE)
```

# we can make some predictions of survival based on different characteristics
```{r}
lpm <- lm(survive ~ child + female + class1 + class2, data = titanic)
print("LPM")
summary(lpm)
lpm_margins <- margins(lpm, at = list(child = 1, female = 1, class1 = 0, class2 = 0))
print(lpm_margins)

#  using probit
summary(pm)
probit_prediction <- pnorm(-1.308744 + 0.6310225 + 1.483612)
print(probit_prediction)

probit_margins <- margins(pm, at = list(child = 1, female = 1, class1 = 0, class2 = 0))
print(probit_margins)

# using logit model
logit_margins <- margins(lm, at = list(child = 1, female = 1, class1 = 0, class2 = 0))
print(logit_margins)

```


