---
title: "Regression Analysis (Econ 0211), Lab6 v2: Hypothesis Testing and Motivating Nonlinear Regression"
author: "Noe Vyizigiro"
date: "April, 2024"
output: html_notebook
---

```{r, include=FALSE}
rm(list = ls())

library(tidyverse)
if(!require("ipumsr")) install.packages("ipumsr")
library(ipumsr)

if(!require("stargazer")) install.packages("stargazer")
library(stargazer)

library(robustbase) #robust library to use for robust regression
library(car) # allow different model tests

if(!require("sandwich")) install.packages("sandwich") # for robust regression
library(sandwich)
if(!require("lmtest")) install.packages("lmtest")
library(lmtest)

if (!require("survey")) install.packages("survey")
library(survey)
if (!require("MASS")) install.packages("MASS")
library(MASS)

```

# comment:
There is a little difference between this data set and the one provided in stata. For example, leisure values are a little different, which translate in a different mean. Weights values are also different.
```{r}
source("/Volumes/middfiles/Classes/Fall23/ECON0211A/Noe_Labs/NV.R")
```

#read data
```{r, include=FALSE}
ddi<-read_ipums_ddi("atus_00031.xml")
atus<-read_ipums_micro(ddi)

ATUS<-read.csv("IPUMS_ATUS_2019.csv")

```

```{r}
#save(atus, file = "atus.csv")

ipums_val_labels(atus$KIDUND18)

columns <-c( "MARST", "SEX")
atus[, columns] <- lapply(columns, function(x) as_factor(atus[[x]]))
columns <-c("AGE", "EARNWEEK_CPS8", "NCHLT5")
atus[, columns] <- lapply(columns, function(x) as.numeric(atus[[x]]))

atus19<-atus%>%
  filter(YEAR==2019)
```
# 2. Generating new Variables
```{r}
atus19<-atus19%>%
  mutate(LTHS =ifelse(EDUC<20, 1,0),
         HS =ifelse((EDUC>=20 & EDUC<=21),1,0),
         SC =ifelse((EDUC>=30 & EDUC<=32), 1,0),
         BA =ifelse(EDUC==40, 1, 0),
         GRAD =ifelse((EDUC>40 & EDUC<=43), 1,0),
         employed =ifelse(EMPSTAT==1, 1, 0),
         male =ifelse(SEX=="Male", 1,0),
         child =ifelse(KIDUND18 ==1, 1, 0))

#atus19
```

# summarize the 10 variables
```{r}
# Selecting the variables of interest
selected_vars <-dplyr::select(atus19,LTHS, HS,SC, BA, GRAD, employed, male,child, Leisure_211, AGE)

# Calculating summary statistics for the selected variables
summary_table <- sapply(selected_vars, function(x) {
  c(
    Obs = sum(!is.na(x)),
    Mean = mean(x, na.rm = TRUE),
    Median = median(x, na.rm = TRUE),
    `Std. dev.` =sd(x, na.rm = TRUE),
    Min = min(x, na.rm = TRUE),
    Max = max(x, na.rm = TRUE)
  )
})

# Print the summary table
summary_df <- as.data.frame(t(summary_table))
print(summary_df)

# or you can get the same results in one line of code by using the nv_sum() function.
nv_sum(atus19, LTHS, HS,SC, BA, GRAD, employed, male,child, Leisure_211, AGE)
```

# Weighted average
```{r}
# Create a survey design object, from survey package
sv_obj <- svydesign(id = ~1, weights = ~WT06, data = atus19) # id: formula specifying cluster id; ~1 indicate that there is no cluster; weights: formula specifying the sampling weights

# Function to calculate weighted mean and standard error
summary_stats <-function(x,y) { # supply function iterate through the selected variables x while using the weights in the sv_object (y)
  c(
    obs =sum(!is.na(x)),
    mean=svymean(x,y, na.rm=TRUE),#calculate the weighted mean
    Std.dev = sqrt(svyvar(x,y, na.rm = TRUE)),# calculate the weighted SD
    min = min(x, na.rm = TRUE),
    max =max(x, na.rm = TRUE)
   # quant=svyquantile(x,y, 0.50, na.rm=TRUE)
  )
}
sum_table<-sapply(selected_vars, summary_stats, sv_obj) # use the sapply function here to the whole created function to create a vector-type data
sum_table
# Convert to a data frame for better presentation
results_df <- as.data.frame(t(sum_table))
# Print the results
print(results_df)

```

# 3 Estimating Regressions
```{r, include=TRUE}
m1<-lm(Leisure_211~HS+ SC+ BA +GRAD, data = atus19)
#summary(m1)

m2<-lm(Leisure_211~HS+SC+ BA+ GRAD, data =atus19, weights = WT06) # non robust regression
#summary(m2)

#robust regression
m3<-lm(Leisure_211~HS+SC+ BA+ GRAD, data =atus19, weights = WT06) 

#rob<-coeftest(m3, vcov = vcovHC(m3, "HC1"))    # robust; HC1 (Stata default)
rob_se3<-sqrt(diag(vcovHC(m3,"HC1"))) #calculate robust standard error
#rob_se3

stargazer( m1,m2,m3,
          se =list(NULL, NULL, rob_se3),
          type = "text",
          title = "Table 1",
          out = "table1.text",
          column.labels = c("Regular", "Weighted", "Robust SE 1"),
          notes = "Significance level"
          
          )
        
```

# 4 Re-estimating the regressions using robust

```{r}
m1<-lm(Leisure_211~HS + SC + BA + GRAD, data =atus19, weights = WT06)
r_se1<-sqrt(diag(vcovHC(m1,"HC1")))

m2<-lm(Leisure_211~HS + SC + BA + GRAD + employed, data =atus19, weights = WT06)
r_se2<-sqrt(diag(vcovHC(m2,"HC1")))

m3<-lm(Leisure_211~HS + SC + BA + GRAD + employed + male, data =atus19, weights = WT06)
r_se3<-sqrt(diag(vcovHC(m3,"HC1")))

m4<-lm(Leisure_211~HS + SC + BA + GRAD + employed + male + child, data =atus19, weights = WT06)
r_se4<-sqrt(diag(vcovHC(m4,"HC1")))

m5<-lm(Leisure_211~HS + SC + BA + GRAD + employed + male + child + AGE, data =atus19, weights = WT06)
r_se5<-sqrt(diag(vcovHC(m5,"HC1")))

stargazer(m1, m2, m3, m4,m5,
          se=list(r_se1, r_se2, r_se3, r_se4, r_se5),
          type = "html",
          title = "Table 2: Weighted Regression with Robust Standard Error.",
          out="table2.html",
          notes.append = TRUE)
```
#functions

```{r}
Reg_moded<-function(model_types, base_formula, ad_var,control_var, data, weights = NULL){
  # initialize the list to store model results
models<-list()
sum<-list()
#create the base formula
base_formula_ob<-as.formula(base_formula)
# iterate over each additional variable
for (i in seq_along(ad_var)){
  formula<-as.formula(paste(deparse(base_formula_ob),"+", ad_var[i]))
  
  # iterate over each model type
  for(model_type in model_types){
    model<-switch (model_type,
                 lm =lm(formula, data=data, weights = weights),
                 lmrob=lmrob(formula, data = data, weights = weights),
                 stop("unsupported model"))
  

  
  model_sum<-summary(model)
  coeff<-coef(model)
  
  models[[paste(model_type, ad_var[i], sep = "_")]] <- model
      sum[[paste(model_type, ad_var[i], sep = "_")]] <- list(
        formula = formula,
        coefficients = coeff
      )
}
}
  "
  models[[i]]<-model
  sum[[i]]<-list(
    formula = formula,
    coefficients=coeff
   #standard_errors = standard_errors
  )
  "
  # Print the results
   # cat("Model", i, ":\n")
  #  cat("Formula:", deparse(formula), "\n")
  #  cat("Coefficients:\n")
   # print(model_sum)
  #  cat("\n")
   
  # fit model with control variables
  for(i in seq_along(control_var)){
    formula<-as.formula(paste(deparse(base_formula_ob),"+", control_var[i]))
    
    #iterate over each model type
    for(model_type in model_types){
    model<-switch (model_type,
                 lm =lm(formula, data=data, weights = weights),
                 lmrob=lmrob(formula, data = data, weights = weights),
                 stop("unsupported model"))
    
      model_sum <- summary(model)
      coeff <- coef(model)
      
      models[[paste(model_type, control_var[i], sep = "_")]] <- model
      sum[[paste(model_type, control_var[i], sep = "_")]] <- list(
        formula = formula,
        coefficients = coeff
      )
  }
  }
   stargazer(models,
          #se =list(NULL, NULL, rob_se3),
          type = "text",
          title = "Table 1",
          out = "table1.txt",
          #column.labels = c("Regular", "Weighted", "Robust SE 1"),
          notes = "Significance level"
          
          )  
  

  return(list(models = models, summaries = sum))
}


base_formula<-"Leisure_211~HS + SC"
ad_var<-c("employed")
control_var<-c("male","child")
model_types<-c("lm", "lmrob")

results<-Reg_moded(model_types,base_formula,ad_var, control_var, atus19, atus19$WT06)


```
#5

```{r}
#b
m1<-lm(Leisure_211~ HS + SC + BA + GRAD + employed + AGE + male + child, data = atus19, weights = WT06)
r_seM1<-sqrt(diag(vcovHC.default(m1, "HC1")))
summary(m1)
r_seM1

t1<-linearHypothesis(m1, "BA =GRAD")
print(t1)
#c
t2<-linearHypothesis(m1, c("HS", "SC", "BA", "GRAD"))
print(t2)

```
#6
```{r}
#a
linearHypothesis(m1, c("child =0", "employed=0"))
linearHypothesis(m1, "child-employed=0") # or linearHypothesis(m1, "child=employed =0)
linearHypothesis(m1, "2*child =employed")

#b
linearHypothesis(m1, c("HS", "SC", "BA", "GRAD", "employed", "AGE", "male", "child "))
```

#7 Age and leisure
```{r}
#c
ggplot(atus19, aes(AGE, Leisure_211))+
  geom_point()+
  labs(
    title = "scatter plot of age and Leisure"
  )
#Scatter plot is not helpful here, we can try something else
sv_ob <- svydesign(id = ~1, weights = ~WT06, data = atus19) 

 sum<-atus19%>%
   group_by(AGE)%>%
   summarise(mean_leisure = weighted.mean(Leisure_211, WT06),
             n=n())
 #sum
 
 #generate a variable that is the average leisure by age
atus19 <-atus19%>%
   group_by(AGE)%>%
   mutate(leisure_byAge = mean(Leisure_211))
  
 ggplot(atus19, aes(AGE, leisure_byAge))+
   geom_line()+
   labs(
     title = "Leisure by age and Age"
   )

 # regress leisure on age
 m1<-lm(Leisure_211~AGE, data = atus19, weights = WT06)
 summary(m1)
 #robust
 rm1<-coeftest(m1,vcov=vcovHC(m1,"HC1")) #robust coefficient and se
 se_m1<-sqrt(diag(vcovHC(m1, "HC1")))

 
 #atus19$pred_leisure<-predict(m1)
 atus19$fitted_values<-fitted(m1)
 
ggplot(atus19, aes(AGE, leisure_byAge))+
  geom_line(aes(color ="Leisure_byAge"))+
  geom_line(aes(y= fitted_values, color ="fitted_values"))+
   labs(
     title = "Predicted values"
   )
```

# The following steps are ones we will explore in class next week ......
```{r}
#d
atus19<-atus19%>%
  mutate(Age2 =AGE^2)

for (i in 2:5){
  new_val_name <- paste0("Age", i)
  atus19[[new_val_name]]<-atus19$AGE^i
  
}
#atus19%>%
 # dplyr::select(AGE, Age2)
```
# Run Regressions 

```{r, include=FALSE}
m1<-lm(Leisure_211~ AGE, weights = WT06, data = atus19)
atus19$pred_yht1<-predict(m1)

m1se<-sqrt(diag(vcovHC.default(m1, "HC1")))

m2<-lm(Leisure_211~ AGE+ Age2, weights = WT06, data = atus19)
atus19$pred_yht2<-predict(m2)

m2se<-sqrt(diag(vcovHC.default(m2, "HC1")))

m3<-lm(Leisure_211~ AGE + Age2 + Age3, weights = WT06, data = atus19)
atus19$pred_yht3<-predict(m3)

m3se<-sqrt(diag(vcovHC.default(m3, "HC1")))

m4<-lm(Leisure_211~ AGE + Age2 + Age3 + Age4, weights = WT06, data = atus19)
atus19$pred_yht4<-predict(m4)

m4se<-sqrt(diag(vcovHC.default(m4,"HC1")))

stargazer(m1, m2, m3, m4,
          se =list(m1se, m2se, m3se, m4se),
          type = "html",
          title = "Table 3.",
          out="table3.html",
          notes.append = TRUE)

```

#plot
```{r}
ggplot(atus19, aes(AGE, leisure_byAge, color = "leisure_byAge"))+
  geom_line()+
  geom_line(aes(y =pred_yht1, color ="pred_yht1"))+
   labs(
     title = "Predicted values"
   )


ggplot(atus19, aes(AGE, leisure_byAge, color = "leisure_byAge"))+
  geom_line()+
  geom_line(aes(y =pred_yht1, color ="pred_yht1"))+
  geom_line(aes(y =pred_yht2, color ="pred_yht2"))+
  geom_line(aes(y =pred_yht3, color ="pred_yht3"))+
  geom_line(aes(y =pred_yht4, color ="pred_yht4"))+
   labs(
     title = "Predicted values"
   )

```

```{r}
ggplot(atus19, aes(AGE, leisure_byAge, linetype = "leisure_byAge", color="leisure_byAge"))+
  geom_line()+
  geom_line(aes(y =pred_yht1, linetype ="linear",color = "linear"))+
  geom_line(aes(y =pred_yht2,linetype ="quadratic",color="quadratic"))+
  geom_line(aes(y =pred_yht3, linetype ="cubic", color="cubic"))+
  geom_line(aes(y =pred_yht4, linetype ="quartic",color="quartic"))+
  scale_linetype_manual(values = c("linear"="dashed", "quadratic"="dotted", "cubic"="twodash", "quartic"="dotdash", "leisure_byAge"="solid"))+
  scale_color_manual(values = c("linear"="blue", "quadratic"="green", "cubic"="purple", "quartic"="red", "leisure_byAge"="black"))+
   labs(
     title = "Predicted values",
     linetype ="Model", # line type will merge with color type to created an annotated, colored legend based on linetype and color type
     color ="Model",
     y="Leisure by Age"
   )+
  theme_minimal()
```


