---
title: "Regression Analysis (Econ 0211), Lab3: Lab 4: Getting Started with Multiple Regressions"
author: "Noe Vyizigiro"
date: "March, 2024"
output:
  html_document:
    theme: cerulean
    toc: TRUE #include table of content
    toc_flaot: TRUE
---
## *Important packages (libraries) for this lab*:
1. tidyverse
2. haven
3. stargazer
4. gt: for saving summary tables
# use the data from stata
```{r}

library(tidyverse)
library(haven) #to help merge
library(stargazer)

if(!require("gt")) install.packages("gt")
library(gt)

rm(list = ls())
```
# Load data
```{r}
usa80_lab4<-read_csv("Ipums_USA_census_1980.csv")
cpi<- read_csv("cpi_inflation_adju.csv")
usa21_lab4<-read_csv("IPUMS_USA_ACS_2021_Lab_4.csv")

# load the nv.R function package
source("/Volumes/middfiles/Classes/Fall23/ECON0211A/Noe_Labs/NV.R") 

```

# 2. Append data
```{r}
usa_earnings<-bind_rows(usa21_lab4, usa80_lab4)%>%
  dplyr::select(-cpi, -cpi_2021_factor, -cpi_adjust_to_2021, -real_earn, -ln_earn, -college) # remove these variables

```

# Tab sex year, sum(earning)
```{r}
tab1<-usa_earnings%>%
  group_by(year, sex)%>%
  summarise(mean_earning=mean(earnings,na.rm=TRUE))%>%
  pivot_wider(names_from = year,
              values_from = mean_earning)%>%
  gt()%>%
  tab_header(
    title = "Average earnings accross years and gender"
  )
tab1
#gtsave(tab1, "aver_earning.html") # save it in html, it can be opened in any format

# or
usa_sex<-usa_earnings%>%
  group_by(year, sex)
nv_sum(usa_sex, earnings, weight = FALSE)%>%
  gt()%>%
  tab_header(
    title = "Average earnings accross years and gender")%>%
  gtsave( "aver_earning.html") # save it in html, it can be opened in any format
```

# 4. Merge usa_earnings with cpi
```{r}
Earn_infra<-merge(usa_earnings, cpi, by="year", all.x = TRUE)

# generate real_earning
Earn_infra<-Earn_infra%>%
  mutate(real_earn = earnings*cpi_adjust_to_2021)

Earn_infra%>%
  group_by(year, sex)%>%
  summarise(mean_earn =mean(real_earn, na.rm=TRUE))%>%
  pivot_wider(names_from = year,
              values_from = mean_earn)
# or
earn_inf<-Earn_infra%>%
  group_by(year, sex)
nv_sum(earn_inf, real_earn, weight = FALSE)
```
# 5. Generate log real_earn
```{r}
Earn_infra<-Earn_infra%>%
  mutate(ln_earn =log(real_earn))

```

# Regressions (fit the models)
```{r}
# Perform linear regression
m1 <- lm(ln_earn ~ female, data = Earn_infra, subset = (year == 1980))
summary(m1)
m2 <- lm(ln_earn ~ female, data = Earn_infra, subset = (year == 2021))
summary(m2)

#create a regression table

```

# 6. Over 40 hrs and weeks worked
```{r, include=FALSE}
Earn_infra%>%
  group_by(wkswork2)%>%
  summarise(mean_earn =mean(real_earn, na.rm=TRUE))
#creating dummy variable

Earn_infra<-Earn_infra%>%
  mutate(Over40Weeks = case_when(wkswork2=="40-47 weeks"~1,
                                 wkswork2=="48-49 weeks"~1,
                                 wkswork2=="50-52 weeks"~1,
                                 TRUE~0))
Earn_infra$uhrswork2<-as.numeric(Earn_infra$uhrswork) # make it a numeric
```


```{r, include=FALSE}
Earn_infra%>%
  dplyr::select(uhrswork, uhrswork2)
```


```{r}
Earn_infra%>%
  group_by(year)%>%
  summarise(mean_hr=mean(uhrswork2, na.rm=TRUE),
            n=n())
```

# Run regressions taking account weeks and hours worked
```{r}
m3<-lm(ln_earn~female, data =Earn_infra, subset =(Over40Weeks ==1 & uhrswork2>=40 & year ==1980 ))
summary(m3)

m4<-lm(ln_earn~female, data =Earn_infra, subset =(Over40Weeks ==1 & uhrswork2>=40 & year ==2021 ))
summary(m4)

```
# 7. Including education

```{r}
# creating a college variable of at least having a college degree
Earn_infra<-Earn_infra%>%
  mutate(college =ifelse(edyears>=16,1,0))

```

# 8. Estimating regression, including college
```{r}
m5<-lm(ln_earn~female+college, data =Earn_infra, subset = (Over40Weeks ==1 & uhrswork2>=40 & year ==1980))
summary(m5)

m6<-lm(ln_earn~female+college, data =Earn_infra, subset = (Over40Weeks ==1 & uhrswork2>=40 & year ==2021))
summary(m6)
 
```

# 9 & 10. Combining all the regressions using stargazer function
```{r, include=TRUE}

stargazer( m1, m2, m3, m4,m5, m6,
          type = "text",
          title = "Table 3. Log earnings on Female for years 1980 and 2021",
          dep.var.caption = "DV: earings in year 1980 and 2021",
          out = "table3.txt",
          column.labels = c("80b","21b","80tb","21tb","80y","21y"),
          notes = "Significance level"
          )

```

### estimating and testing correlations
```{r}
cor(Earn_infra$female,Earn_infra$college, method = "pearson")
cor.test(Earn_infra$female,Earn_infra$college, method = "pearson")

```

# 11. Creating race dummies
```{r}
Earn_infra%>%
  group_by(race)%>%
  summarise(mean = mean(real_earn))

Earn_infra<-Earn_infra%>%
  mutate(white_nh=ifelse(race=="white" & hispan=="not hispanic", 1,0),
         black_nh=ifelse(race=="black/african american" & hispan =="not hispanic",1, 0),
         hispan_nh=ifelse(hispan!="not hispanic", 1, 0),
         other_nh=ifelse(race!="white" & race !="black/african american" & hispan =="not hispanic",1,0))

nv_sum(Earn_infra, white_nh, black_nh, hispan_nh, other_nh, weight = FALSE)  
```



