---
title: "Regression Analysis (Econ 0211), Lab2 :Getting Ready for Regression"
author: "Noe Vyizigiro"
date: "March, 2024"
output: html_notebook
---

```{r setup,}
library(ipumsr)
library(tidyverse)
library(stringr)
library(sf) #maps
library(purrr)
library(haven)
install.packages("RStudio")

rm(list=ls())
```

This version uses data downloaded directly from IPUMS
## Load the data
```{r}
ddi <-read_ipums_ddi("usa_00001.xml")
usa_21d<-read_ipums_micro(ddi) #point to the code book to read the file: data file is dump, but the code book is smart,it has everything we need.
#as_factor(usa_bdi) #to see the labels only for the variables that are labled
#zap_labels(usa_bdi) # to see only the numeric values associated with the labels 

#load nv.R function package
source("/Volumes/middfiles/Classes/Fall23/ECON0211A/Noe_Labs/NV.R") 

```
```{r}
names(ddi) # the name element of the ddi objects; it has a file name, which is the data file that we have to download and save in the same location with ddi book code
ddi$var_info
```
```{r}
names(usa_21d) # name variables
ipums_var_label(ddi, EDUC) # view the variable's label
ipums_var_desc(ddi, EDUC) # view longer description of a variable
ipums_view(ddi)
usa_21d

```

```{r}
ipums_val_labels(usa_21d$EDUC) # look at a variable's values and their labels
```

## making some variables factors
```{r}
columns <-c("EDUC", "EDUCD", "SEX")
usa_21d[, columns] <- lapply(columns, function(x) as_factor(usa_21d[[x]]))

columns0 <-c("AGE", "INCWAGE", "INCEARN")
usa_21d[, columns0] <- lapply(columns0, function(x) as.numeric(usa_21d[[x]]))

usa_21d

```


## 4. Looking at the earning variables
```{r}
min(usa_21d$INCWAGE, na.rm = TRUE)
max(usa_21d$INCWAGE, na.rm = TRUE)
mean(usa_21d$INCWAGE, na.rm = TRUE)
median(usa_21d$INCWAGE, na.rm = TRUE)
sd(usa_21d$INCWAGE, na.rm = TRUE)
var(usa_21d$INCWAGE, na.rm = TRUE)

#or
#summarize INCWAGE AND INCEARN
nv_sum(usa_21d, INCWAGE, INCEARN, weight = FALSE) # using nv_sum() function


```

### replace 0s to nas and drop na to avoid missing values
```{r}
usa_21d$INCWAGE[usa_21d$INCWAGE ==0]<-NA # replacing zeros with NAs
usa_21d$INCWAGE[usa_21d$INCWAGE ==999999]<-NA


#create a new dataset without missing values
usa_21d0<-usa_21d%>%
  na.omit()
summary(usa_21d$INCWAGE)
```

```{r}
table(usa_21d0$EDUC)
table(usa_21d0$EDUCD)

```

```{r}

usa_21d1<-usa_21d0%>%
  mutate(edyears = as.numeric(EDUCD)) #create a new variable to observe numerical labels of educD

#usa_21d1
```

```{r}
usa_21d1%>%
  group_by(EDUCD, edyears)%>%
  summarise(mean = mean(INCWAGE))



```

## 6.generating a new variables "edyears" with numbers of years of education
```{r}
usa_21d2<-usa_21d1%>%
  mutate(edyears =recode(edyears, '2' ='0', '3'='0', '5'='0', '6' ='0', '8'='1', '9'='2', '10'='3', '11'='4', '14'='5', '15'='6', '17'='7', '18' ='8', '19'='9', '20'='10', '21'='11', '23' ='11', '25'='12','26'='12', '27'='13', '29'='13', '31'='14', '36'='16', '41' ='18', '42' ='18', '43'='20' ),
        edyears = as.numeric(edyears) )
```
### check if the matching was successful
```{r}
usa_21d2%>%
  group_by(EDUCD, edyears)%>%
  summarise(mean= mean(INCWAGE))
```

## 7. Grouping by edyears to observe income wage across years of education.
```{r}
# grouping by educ to observe income wage across years of education.
usa_21d2%>%
  group_by(edyears)%>%
  summarise(mean = mean(INCWAGE))

# creating and using an dataset with age between 25 and 60
usa_21d3<-usa_21d2%>%
  filter(AGE>=25 & AGE<=60)

usa_21d3<-usa_21d3%>%
  group_by(edyears)
  #summarise(mean_wage =mean(INCWAGE))
nv_sum(usa_21d3, INCWAGE, weight = FALSE)
```

## 8. Compare average earnings for college and non-college grads
```{r}
# create a dummy variable college equal to 1 if educ is >=16 and zero otherwise
usa_21d3%>%
  mutate(college = ifelse(edyears>=16,1,0))%>%
  group_by(college)%>%
  summarise(mean_wage = mean(INCWAGE))
```

## 9. Graph
```{r}
usa_21d3%>%
  ggplot(aes(edyears, INCWAGE))+
           geom_point()+
  labs(
    title = "Scatter Plot of Education vs. Earnings",
    subtitle="2021 American Community Survey",
    x="Years of education",y="wage($)"
  )
```

```{r}
usa_21d3%>%
  filter(AGE>=25 & AGE<=60)%>%
  ggplot(aes(edyears, INCWAGE))+
           geom_point()+
  geom_smooth(method=lm, se =FALSE)+
  labs(
    title = "Scatter Plot of Education vs. Earnings",
    subtitle="2021 American Community Survey",
    x="Years of education",y="wage($)"
  )
 
```
### How to mine data from ipums compatible with r:
Assuming one has already an ipums account:
1. Go to ipums and select your sample and all the variables 
2. Hit "extract dataset"
3. hit " submit extract" dataset in ".dat-format
4. When the dataset is ready, download the dat format on your laptop (it comes as a compressed dat.gz file)
5. download the ddi file(the "code book") by double clicking on it and click on "save as.."
6. save both files in the same directory
7. read the ddi file in r : ddi<-read_ipums_ddi(data.xml)
8. create a dataset to be used for your analysis using the ddi file loaded: data<-read_micro_ipums(ddi)
9. then, you're good to go with your data cleaning and analysis.


### Saving the cleaned data set
```{r}
write.csv(usa_21d3, file = "usa21_earnings.csv") # this is to save the cleaned data to the directory to be used later for lab 3
#read.csv("new_data.csv")
```


