---
title: "Regression Analysis (Econ 0211), Lab2 :Getting Ready for Regression"
author: "Noe Vyizigiro"
date: "March, 2024"
output:
  html_document:
    theme: cerulean
    toc: TRUE #include table of content
    toc_flaot: TRUE
---
## *Important packages (libraries) for this lab*:
1. tidyverse: Contain libraries such as dplyr and ggplot, which are instrumental here
2. purrr

# load necessary libraries
```{r setup, }

library(tidyverse)
if(!require("purrr")) install.packages("purrr")
library(purrr) #this will help in summarizing the variable

rm(list = ls()) #clear the environment
```
This version uses data converted using Stata
# Load the data
```{r}
usa21<-read_csv("IPUMS_USA_2021_ACS.csv")

# make some variables numeric
columns <-c("incwage", "incearn", "year")
usa21[, columns] <- lapply(columns, function(x) as.numeric(usa21[[x]]))

```

# 4. Looking at the earning variables

*Creating a function that can take in multiple arguments(variable) and return a table of the variables with their summary statistics. We use function enquos to capture multiple variables as a list of quosures. We use map_dfr function from the package purrr to iterate over each variable, calculate summary statistics, and combine the results into a single data frame.*
```{r}
nv_sum<-function(datax,...){
  vars<-enquos(...) # capture multiple variables and assign them to the variable vars
  map_dfr(vars, function(varx){ # iterate over the variables in vars to compute summary statistics
  var_name<-as_label(varx)
  datax%>%
    summarise(variable =var_name,
              min =min({{varx}}, na.rm = TRUE),
              mean =mean({{varx}}, na.rm=TRUE),
              median =median(!!varx, na.rm =TRUE),
              st.dev =sd(!!varx, na.rm =TRUE),
              max=max(!!varx, na.rm = TRUE),
              n =n()
    ) 
  })
}
# varx can be either between {{}} or start with !! to ensure that the column name passed to the function is correctly referenced within summarise
nv_sum(usa21, incwage, incearn)


```

```{r}
filtered_data<-usa21%>%
  filter(age>=25 & age<=60)

nv_sum(filtered_data, incwage) # call the nv_sum function created already
```

We only want to look at people with positive earnings (for now) so set all others to missing 
set our earnings variable to missing if respondent has no employee income and/or is not in university for incomewage
```{r}
usa21$incwage[usa21$incwage==0]<-NA #replacing zeros in the incwage variable with NAs
usa21_0<-usa21 %>% 
  na.omit() # get rid of all missing values 

```

# 6. Look for an education variable and recorde
```{r}
table(usa21_0$educ)
table(usa21_0$educd)
 #or
usa21_0%>%
  count(educ)

```
## Generate numeric values of the values in variable educd

```{r}
usa21_0<-usa21_0%>%
  mutate(educCode=with(usa21_0, match(educd, unique(educd))))
```
or the same thing can be achieved by:
usa21_0$educCode = as.integer(factor(usa21_0$educd, levels = unique(usa21_0$educd)))

## Summarize educd with educCode to see the matches
```{r}

# or we can group by educd and use the new dataset as an argument in our function nv_sum().
grouped_data<-usa21_0%>%
  group_by(educd)

nv_sum(grouped_data, educCode)

```


## Generating a new variables "edyears" with numbers of years of education
```{r}
usa21_1<- usa21_0%>% 
   mutate(edyearF =factor(educCode, c(2, 13, 8, 11, 9,7,24,3,6,22, 18, 19, 21, 16, 5, 17, 15, 20, 10, 14, 23, 12, 1, 4 ), 
 labels=c('13','11','14','16','20','12','1','10','11','2','3','4','5','6','7','8','9','0','18','0','0','18','12','13')),
 edyears = as.numeric(as.character(edyearF)))
 
usa21_1

```

## Check if the match was successful
```{r}
usa21_1%>%
 group_by(educd) %>% 
  summarise(means = mean(edyears))
```

# 7. Relationship between wage and years of education
```{r}
edyear_usa21_1<-usa21_1%>%
  group_by(edyears)
nv_sum(edyear_usa21_1, incwage) # call the function nv_sum()
 
```

```{r}
gredyear_usa21_1<-usa21_1%>%
  group_by(edyears)%>%
  filter(age>=25 & age<=60) %>% 
  summarise(mean_wage=mean(incwage) )
```

# 8. Compare average earnings for college and non-college grads
```{r}
usa21_1%>%
  mutate(college_grads =ifelse(edyears>=16,1,0))%>%
  group_by(college_grads)%>%
  filter(age>=25 & age<=60)%>%
  summarise(mean_wage=mean(incwage) )
```

# 9.Graphs
```{r}
usa21_1%>%
  filter(age>=25 & age<=60)%>%
  ggplot(aes(edyears, incwage))+
           geom_point()+
  labs(
    title = "Scatter Plot of Education vs. Earnings",
    subtitle="2021 American Community Survey",
    x="Years of education",y="wage($)"
  )
  
```

```{r}

usa21_1%>%
  filter(age>=25 & age<=60)%>%
  ggplot(aes(edyears, incwage))+
           geom_point()+
  geom_smooth(method=lm, se =FALSE)+
  labs(
    title = "Scatter Plot of Education vs. Earnings",
    subtitle="2021 American Community Survey",
    x="Years of education",y="wage($)"
  )
  
```








