---
title: "Regression Analysis (Econ 0211), Lab12: FIGURES AND TABLES"
author: "Noe Vyizigiro"
date: "May, 2024"
output:
  html_document:
    theme: cerulean
    toc: TRUE #include table of content
    toc_flaot: TRUE
---
# *Important Packages/Libraries for this Lab:*


```{r}
options(repos = c(CRAN = "https://cloud.r-project.org"))

library(tidyverse)
if(!require("sandwich")) install.packages("sandwich")
library(sandwich)
library(lmtest)
if(!require("dotwhisker")) install.packages("dotwhisker")
library(dotwhisker) # coefplots

if(!require("broom")) install.packages("broom")
library(broom) # coefplots

if(!require("margins")) install.packages("margins")
library(margins) # coefplots

rm(list = ls())

source("/Volumes/middfiles/Classes/Fall23/ECON0211A/Noe_Labs/NV.R") #load NV.R to access nv_sum() function for quick summary statistic tables
```

# Read Data
```{r}
usa77<-read_csv("IPUMS_USA_1970_2017.csv")
usa77
```
```{r}

usa77 <-usa77%>%
  mutate(codeduc=with(usa77, match(educ, unique(educ))))

usa77%>%
  count(empstat)

 
```
# Creating education categories using the numerical values in educnum
```{r}
usa77<-usa77%>%
  mutate(edcat =case_when(codeduc== 1 |codeduc== 3 |codeduc== 4 |codeduc== 5 | codeduc==10 |codeduc== 12 ~"Less than HS",
                          codeduc ==2 ~ "High School",
                          codeduc ==7 |codeduc==8 |codeduc==11 ~"Some College",
                          codeduc ==6~ "Bachelors",
                          codeduc ==9~"Graduate",
                          TRUE~"N/A"),
         inlf=ifelse(empstat!="not in labor force",1,0))
usa77%>%
  count(edcat, educ, edcat, empstat, inlf)
```

# Including weights
```{r}
sex_inlf<- function(data, gender){
  data %>%
  filter(sex=={{gender}})%>%
  group_by(year, edcat)%>%
  summarize(mean_inlf = sum(inlf * perwt) / sum(perwt), .groups = 'drop')%>%  # weighted mean of inlf
  pivot_wider(names_from = edcat,
              values_from = mean_inlf)
}
 
sex_inlf(usa77, "male") # male
sex_inlf(usa77,"female") # female
```

# A Small Summary stats Table
```{r}
usa77<-usa77%>%
  mutate(LTHS=ifelse(edcat=="Less than HS",1,0),
         HS =ifelse(edcat =="High School",1,0),
         SC =ifelse(edcat=="Some College",1,0 ),
         BA =ifelse(edcat =="Bachelors", 1, 0),
         GRAD =ifelse(edcat =="Graduate", 1, 0)
         )
```

# summary tables
```{r}
nv_sum(usa77, LTHS, HS, SC, BA, GRAD, inlf, weight = TRUE)

nv_sum(usa77, LTHS, HS, SC, BA, GRAD, inlf, weight = TRUE, condition = quo(year==1970))
nv_sum(usa77, LTHS, HS, SC, BA, GRAD, inlf, weight = TRUE, condition = quo(year==2017))
```

# Making summary tables by grouping
```{r}
nv_sum(usa77%>%group_by(year), LTHS, HS, SC, BA, GRAD, inlf, weight = TRUE) %>%
  dplyr::select(year, variable, mean)%>%
  pivot_wider(names_from =variable ,
              values_from = mean)

nv_sum(usa77%>%
         group_by(year, sex),
       LTHS, HS, SC, BA, GRAD, inlf, weight = TRUE) %>%
  dplyr::select(year, sex, variable, mean)%>%
  pivot_wider(names_from = variable,
              values_from = mean)
```
# Create new Variables
```{r}
usa77$uhrswork<-as.numeric(usa77$uhrswork)


usa77$edcat_num<-as.numeric(factor(usa77$edcat, levels = c("Less than HS", "High School", "Some College", "Bachelors", "Graduate"))) # creating edcat_num,  based on edcat
usa77$edcat_num<-as.factor(usa77$edcat_num) # converting it to factor
ed<-as.data.frame(model.matrix(~edcat_num-1, data = usa77))
colnames(ed)<-paste0("ed", seq_along(colnames(ed)))
usa77_2<-cbind(usa77, ed)


usa77_2<-usa77_2%>%
  mutate(codehrs=with(usa77, match(hrswork2, unique(hrswork2))),
         hrover40_2 =ifelse(hrswork2=="40 hours" |hrswork2=="41-48 hours"|hrswork2=="49-59 hours"|hrswork2=="60+ hours",1,0),
         wkover40=ifelse(wkswork2=="40-47 weeks"| wkswork2=="48-49 weeks"|wkswork2=="50-52 weeks",1,0),
         fulltime_year =case_when(uhrswork>=40 & wkover40==1~ "1",
                                  hrover40_2==1 & year=="1970" & wkover40==1~ "1",
                                  hrover40_2==0 & year=="1970" & wkover40==1~ "0",
                                  TRUE~"0"),
         ln_wage =ifelse(incwage>0, log(incwage),NA),
         male =ifelse(sex=="male",1,0),
         age2 = age^2
         
  )
 
#usa77_2%>%
 # count( edcat_num, ln_wage)
```

# try loops
```{r}
for(i in 1:10){
   y<-i*2
   print(y)
}


# alternatives of the loops: apply family functions:
# apply(), lapply(), tapply(), sapply(), vapply(), and mapply()

y<-lapply(1:10, function(x){
  x*2
})
y

# to get results as a list
y<-sapply(1:10, function(x){
  x*2
})
y
 


```


# regression results
```{r}
list_year<-c(1970, 1980, 1990, 2000, 2010, 2017)
for(i in list_year)(
  print(i)
)
```

```{r}
list_year<-c(1970, 1980, 1990, 2000, 2010, 2017)
#ed_list<-c("ed2", "ed3", "ed4","ed5")
results<-list()
results2<-list()
graph<-list()
graph2<-list()

for(i in list_year){
  m<-lm(ln_wage~male, subset = (fulltime_year ==1 & year==i), weights = perwt, data = usa77_2)
  rob_se <- coeftest(m, vcov = vcovHC(m, type = "HC1")) #extract st.error 
  results[[as.character(i)]] <- list(model = m, robust_se = rob_se) #put the models and their robust St.Errors in the list results

  graph[[as.character(i)]]<-m # extract coefficient to plot later
  
  m2<-lm(ln_wage~male+ed2+ed3+ed4+ed5+age+age2, subset = (fulltime_year==1 & year ==i), weights = perwt, data = usa77_2)
  rob_se2 <- coeftest(m2, vcov = vcovHC(m2, type = "HC1")) #extract st.error 
  results2[[as.character(i)]]<-list(model=m2, robust_se = rob_se2)
  
   graph2[[as.character(i)]]<-m2 # extract coefficient to plot later

}
all_models <- c(lapply(results, function(x) x$model), lapply(results2, function(x) x$model)) #combine the models from the model results using x$models (model =key used to store models in the list)
all_se <- c(lapply(results, function(x) x$robust_se[, "Std. Error"]), lapply(results2, function(x) x$robust_se[, "Std. Error"])) #combine rob_se from the model results using x$robust_se (robust_se =key used to store the rob_se in the list)


 stargazer::stargazer(all_models,
                       se=all_se,
                       type = "text",
                       out = "table1.text",
                       title = "Wages",
                       column.labels = rep(as.character(list_year),2), #duplicate labels
                       notes.append = TRUE)

```

# ploting coefficients
```{r}
# 1st way: Using dwplot function
dwplot(graph)+
  labs(
     title="Evolution of the US Gender Gap",
     x="Coefficients"
  )

  
# 2nd method: using ggplot, which allows including more feature and chart manipulation
coef_data <- bind_rows(
  lapply(names(graph2), function(year) {
    tidy(graph2[[year]], conf.int = TRUE) %>%
      filter(term == "male") %>%
      mutate(model = paste0("Model_", year),  # Add a model identifier
             year = year)
  })
)

ggplot(coef_data, 
       aes( x = year,y = estimate, color = year)) +
  geom_point(size =1)+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width=0.2)+
  theme_minimal() +
  labs(title = "Evolution of the US Gender Gap",
       x= "Male",
       y = "Coefficient Estimate") 
  
```

# creating race dummies
```{r}
usa77_2%>%
  group_by(hispan)%>%
  summarise(mean = mean(ln_wage, na.rm = TRUE))

usa77_2<-usa77_2%>%
  mutate(white_nh=ifelse(race=="white" & hispan=="not hispanic", 1,0),
         black_nh=ifelse(race=="black/african american/negro" & hispan =="not hispanic",1, 0),
         hispan_nh=ifelse(hispan!="not hispanic", 1, 0),
         other_nh=ifelse(race!="white" & race !="black/african american/negro" & hispan =="not hispanic",1,0))

#check the proportion of each group
nv_sum(usa77_2, white_nh, black_nh, hispan_nh, other_nh, weight = FALSE)
```


```{r}
usa77_2<-usa77_2%>%
  mutate(racecat=case_when(white_nh==1~"white_nh",
                           black_nh==1~"black_nh",
                           other_nh==1~"other_nh",
                           hispan_nh==1~"hispanic"),
         college =ifelse(edcat=="Bachelors" | edcat=="Graduate", 1,0))
```


```{r eval=FALSE}
usa77_2%>%
  count(edcat)

usa77_2%>%
  dplyr::select(college ,edcat)
```


# Ploting predicted lines for each race category
```{r}
mod<-lm(ln_wage~college*racecat, subset = (year==2017), data=usa77_2)
summary(mod)
margins_mod <- margins(mod, variables = "college", at = list(racecat = unique(usa77_2$racecat)))
margins_df <- as.data.frame(margins_mod)

ggplot(margins_df, aes(x = factor(college), y = fitted, color = factor(racecat))) +
  geom_point() +
  #geom_line(aes(group = racecat)) +
  labs(x = "college", y = "Linear Prediction",
       color = "Treated",
       title = "Predicted Ln(wages) by BA status and Race",
       subtitle = "US ACS 2017") +
  theme_minimal()

```

