---
title: "Regression Analysis (Econ 0211), Lab9: Working with Pannel Data"
author: "Noe Vyizigiro"
date: "May, 2024"
output:
  html_document:
    theme: cerulean
    toc: TRUE #include table of content
    toc_flaot: TRUE
---
# *Important packages/libraries for this lab:*

1. tidyverse
2. skimr
3. psych
4. sandwich
5. plm
6. stargazer
7. broom
8. car
9.lmtest
```{r}
options(repos = c(CRAN = "https://cloud.r-project.org"))

#clear the environment
rm(list = ls())

library(tidyverse)

if(!require("skimr")) install.packages("skimr")
library(skimr)# quickly skim the dataset

if(!require("psych")) install.packages("psych")
library(psych) # comprehensive summary table

if(!require("sandwich")) install.packages("sandwich")
library(sandwich) # for robust SE

if(!require("plm")) install.packages("plm")
library(plm) # panel linear model

if(!require("startgazer")) install.packages("stargazer")
library(stargazer) 

if(!require("broom")) install.packages("broom")
library(broom)

if(!require("car")) install.packages("car")
library(car) # F-test and other model tests

if(!require("lmtest")) install.packages("lmtest")
library(lmtest)



```

# Load Data
```{r}
inc_dem<-read_csv("income_democracy.csv")
source("/Volumes/middfiles/Classes/Fall23/ECON0211A/Noe_Labs/NV.R") #load NV.R to access nv_sum() function for quick summary statistic table
```

# 2. Investigate the Data
```{r}
skim(inc_dem) # help to quickly skim the dataset

describe(inc_dem) # provide an comprehensive summary table

nv_sum(inc_dem, dem_ind, log_gdppc, log_pop, weight = FALSE)

```
# 3. Investigating dem_ind variable
```{r}
sum_dem_ind<-inc_dem%>%
  group_by(country)
nv_sum(sum_dem_ind, dem_ind, weight = FALSE)

```

# 4. Generate variables for averages by country 
```{r}
# sorting/arranging the data
inc_dem<-inc_dem%>%
  arrange(year,country)%>%
  arrange(country, year)%>%
  mutate(dem_average = mean(dem_ind, na.rm = TRUE))

  
# generate a variable that is the mean democracy index for each country 
inc_dem<-inc_dem%>%
  group_by(country)%>%
  mutate(dem_c_avg =mean(dem_ind, na.rm = TRUE))

#generate a variable that is the mean gdppc for each country 
inc_dem<-inc_dem%>%
  group_by(country)%>%
  mutate(gdp_c_avg = mean(log_gdppc, na.rm = TRUE))
 
# quick look at the variables created
inc_dem%>%
  dplyr::select(country, dem_average, dem_c_avg,gdp_c_avg)

```

# 5. Graph: Relatinship between the natural log of GDP per capita and the democracy index.
```{r}

ggplot(inc_dem, aes(x=gdp_c_avg,y=dem_c_avg))+
  geom_point()+
  geom_smooth(method = "lm", se=TRUE, color="red")+
  labs(
    title = "Democracy and Income",
    subtitle = "average 1960-2000",
    x ="Log GDP per capita",
    y = "Democracy Index"
  )
filtered_graph<-inc_dem%>%
  filter(year==1990)%>%
  ggplot(aes(x=log_gdppc,y=dem_ind))+
    geom_point()+
    geom_smooth(method = "lm", se=TRUE, color="blue")+
    labs(
      title = "Democracy and Income",
      subtitle = "1990",
      x ="Log GDP per capita",
      y = "Democracy Index"
    )
filtered_graph
```

# 6. Estimating Simple Robust Regression Model
```{r}

rob <- lm(dem_ind ~ log_gdppc, data = inc_dem)
# Summarize the model
summary(rob)
rob_se<-sqrt(diag(vcovHC(rob,type="HC1"))) #calculate robust standard error
rob_se

```

# 7. Estimating Clustered Regression Model (to avoid autocorrelation)
```{r}
clust_m<-lm(dem_ind~log_gdppc, data = inc_dem)

clu_se<-sqrt(diag(vcovCL(clust_m, cluster = ~country)))


stargazer(rob,clust_m,
          se=list(rob_se, clu_se),
          type = "text",
          out = "robust.html",
          title = "Robust Model",
          notes.append = TRUE)

```

# 9. Let's tell R we have panel data!
```{r}
pinc_dem<-plm::pdata.frame(inc_dem,index= c("country", "year"), drop.index =FALSE, row.names = FALSE) #tell R that we have a panel data; year becomes a factor
#pinc_dem

# nest the variables (country and country id with the numerical series created)
bpinc_dem<-pinc_dem%>%
  complete(nesting(country))
bpinc_dem



```

```{r}
#Here is some nifty code to make new variables that are the difference between dem_ind and log_gdppc between 1970 and 1995
#create lagged values
 bpinc_dem<-bpinc_dem %>%
  dplyr:: mutate(
    dem_lag_7095 = ifelse(year==1995,lag(pinc_dem$dem_ind, 25),NA),
    
    gdp_lag_7095 =ifelse(year==1995, lag(pinc_dem$log_gdppc, 25),NA)
  )
  

# Create differences
bpinc_dem<-bpinc_dem %>%
 dplyr:: mutate(
    dem_diff_7095 = (dem_ind - dem_lag_7095),
    gdp_diff_7095 = ( log_gdppc - gdp_lag_7095)
  ) 
```

```{r}
# view variables created
selected_data <- bpinc_dem %>%
  dplyr::select(country, year, dem_ind,dem_lag_7095, log_gdppc, gdp_lag_7095 ) #call select directly from dplyr package to avoid conflict with other packages

# Print the selected data to check the result
print(selected_data)
```

# Plot
```{r}
pinc_dem_graph <- bpinc_dem %>%
  filter(!is.na(gdp_diff_7095) & !is.na(dem_diff_7095))

#str(pinc_dem_graph)  #check structure of the entire dataset


ggplot(pinc_dem_graph, aes(x = gdp_diff_7095, y = dem_diff_7095)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)+
   labs(
    title = "Changes in Democracy and Income",
    subtitle = "1970 to 1995",
    x ="Changes Log GDP per capita",
    y = "Changes in Democracy Index"
  )

#pinc_dem_graph

```
# using fitted values
```{r}
# We can do the same graph using the fitted value
plm_model<-lm(dem_diff_7095~gdp_diff_7095, data = pinc_dem_graph)
summary(plm_model)

pinc_dem_graph$yhat<-fitted(plm_model)

ggplot(pinc_dem_graph, aes(x=gdp_diff_7095, y=dem_diff_7095))+
  geom_point()+
  geom_line(aes(y=yhat), color="red")+ # add a smoothed conditional mean
  #geom_smooth(method = "lm", se=FALSE, color="blue")+
  labs(
    title = "Changes in Democracy and Income",
    subtitle = "1970 to 1995",
    x ="Changes Log GDP per capita",
    y = "Changes in Democracy Index"
  )

```

# 10. Start Estimating fixed Effect Regression
```{r}
# get rid of na's
bpinc_dem<-bpinc_dem%>%
  filter(!log_gdppc=="NA")

#pinc_dem

#creating country_id dummies
bpinc_dem <- cbind(bpinc_dem, model.matrix(~ country- 1, data = bpinc_dem)) #creating new country dummies starting or rearranging the country series
#bpinc_dem

```

```{r}

# estimating regressions
dum<-lm(dem_ind~log_gdppc + country-1, data = bpinc_dem)
#summary(dum)
cof<-coef(dum)["log_gdppc"]
#cof
du_se<-sqrt(diag(vcovCL(dum, cluster = ~country)))
#du_se

fe<-plm(dem_ind~log_gdppc, model="within",index = "country", data = bpinc_dem)
#summary(fe)
fe_se<-sqrt(diag(vcovSCC(fe, cluster = "group", type = "HC1")))
#fe_se
coeftest(fe, vcov = vcovSCC(fe,cluster = "group", type = "HC1")) #extract clustered robust coefficients.

stargazer(rob,clust_m, dum,fe,
          se=list(rob_se, clu_se, du_se, fe_se),
          omit = "^country",
          type = "text",
          out = "robust.html",
          title = "Robust Model",
          notes.append = TRUE)

coef
```

# 11. Including time fix effects

Note:
*For cross-sectional data(many subject at a single period of time) with heteroskedasticity, use vcovHC: Corrects for heteroskedasticity*
*For panel data with clustering (e.g., individuals observed over time), use vcovCL: Corrects for within-cluster correlation*
*For panel data with cross-sectional and serial correlation, use vcovSCC: Corrects for both heteroskedasticity and serial correlation*
*vcovCL: Mostly for clustering with linear model (lm)*
*vcovSCC: clustering with panel linear model (plm)*
```{r}
# creating time dummies
dpinc_dem<-cbind(bpinc_dem, model.matrix(~year-1, data = bpinc_dem))
#dpinc_dem

yfe<-plm(dem_ind~log_gdppc+ year-1, index="country", data = bpinc_dem) # fixed effect regression with time dummies; index: country
#summary(yfe)

yfey<-plm(dem_ind~log_gdppc+ year-1, index="year", data = bpinc_dem) # fixed effect regression with time dummies; index: year
summary(yfey)

yfe_se<-sqrt(diag(vcovSCC(yfe, type="HC1",cluster = "group")))
#yfe_se

stargazer(rob,clust_m, dum,fe,yfe,
          se=list(rob_se, clu_se, du_se,fe_se, yfe_se),
          omit = "^country",
          type = "text",
          out = "robust.html",
          title = "Robust Model",
          notes.append = TRUE)
# f-test

test<-linearHypothesis(yfe, c("year1960","year1965","year1970","year1975","year1980","year1985" ,"year1990","year1995"))
test

```

# 12. Control for more variables
```{r}
yfe2<-plm(dem_ind~log_gdppc+ log_pop+educ+age_2+age_3+age_4+age_5+year-1, index="country", data = bpinc_dem)
#summary(yfe2)

yfe_se2<-sqrt(diag(vcovSCC(yfe2, type="HC1",cluster = "group")))
#yfe_se

stargazer(rob,clust_m, dum,fe,yfe,yfe2,
          se=list(rob_se, clu_se, du_se,fe_se, yfe_se, yfe_se2),
          omit = "^country",
          type = "text",
          out = "robust.html",
          title = "Robust Model",
          column.labels = c("ro", "cl", "du","fe","yfe","y"),
          notes.append = TRUE)

test1<-linearHypothesis(yfe2, c("age_2","age_3","age_4","age_5"))
test2<-linearHypothesis(yfe2, c("age_2","age_3","age_4","age_5","educ", "log_pop"))
test1
test2
```

# Bonus: Making a visualization of the results
```{r}
#creating a list of our models
M_list<-list(rob,clust_m, dum,fe,yfe,yfe2)
#extract coefficients and CI
coef_extr<-lapply(M_list, function(model){
  coefs<-tidy(model) #from the broom library
  conf_int<-confint(model)
  coefs<-cbind(coefs, conf_int) #combine coefficient and confidence intervals
  coefs
})
"
for(i in M_list){
  coefs<-tidy(i) #from the broom library
  conf_int<-confint(i)
  coefs<-cbind(coefs, conf_int)
}
"
#apply that to all the models using the function created
#model_coef<-lapply(M_list, coef_extr)
for (i in seq_along(coef_extr)){
  coef_extr[[i]]$model<-paste("Model", i)
}

# combine the coefficients in a single dataframe
coef_df<-bind_rows(coef_extr)
#coef_df
# get the coefficients of interest
coef_df<-coef_df%>%
  filter(term== "log_gdppc")

#rename the columns for convenience
colnames(coef_df) <- c("term", "estimate", "std.error", "statistic", "p.value", "conf.low", "conf.high", "model")

models=c("basic", "cluster", "dummies", "fixed_effects", "time_fe", "fe_controls") #rename the models to match the colors

ggplot(coef_df, aes(x=model, y =estimate, color=models))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width=0.2)+
  geom_hline(yintercept = 0,linetype ="dashed", color="red")+ # add a horizontal dashed line
  labs(title = "Coefficients of log_gdppc Across Models",
       x = "Coefficients on Log_gdppc",
       y = "Coefficient Estimate",
       color="Model") +
   scale_color_manual(values = c("basic" = "blue", "cluster" = "green", "dummies" = "red", "fixed_effects"="yellow", "time_fe"="black", "fe_controls"="purple")) +
  theme_minimal()
```


